The effects of uncertain context inference on motor adaptation

### 2022-04-26 adaptation/article/article.tex
First draft for post-Heald version ready

\abstract{Abstract}
Humans have been shown to adapt their movements when a sudden change to the dynamics of the environment is introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. In this work, we introduce a computational model to explain the behavioral phenomena effected by uncertain contextual information. Specifically, we present a hierarchical model for motor adaptation based on exact Bayesian inference. This model explicitly takes into account contextual information and how the dynamics of context inference affect adaptation and action selection. We show how the proposed model provides a unifying explanation for four different experimentally-established phenomena: (i) effects of sensory cues and proprioceptive information on switching between tasks, (ii) the effects of previously-learned adaptations on switching between tasks, (iii) the effects of training history on behavior in new contexts, in addition to (iv) the well-studied savings, de-adaptation and spontaneous recovery.

Humans have been shown to adapt their movements when sudden or gradual changes to the dynamics of the environment are introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains explicit components of context inference and Bayesian motor adaptation. This model was used to show the effects of context inference on learning rates across different experiments. We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go further than previously shown. Here, we used the COIN model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation. Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from imprecise context inference.

@@

In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in motor adaptation \citep{Heald_Contextual_2021}, as well as in other fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}. We build on models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019} by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In the proposed model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.

This work expands on the work by \cite{Heald_Contextual_2021}, who showed that context inference modulates learning rate in cases of unstable contexts. To do this, we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters.

@@

\section{Results}
We present a motor adaptation model in which an existing model of the environment is updated based on error signals produced by the need of adaptation. We expanded on existing models \citep{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019} by adding an explicit component of context inference. As we will show, context inference guides the selection of the adequate forward model, its updating and the sampling of actions.

We simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference. We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.

Before presenting our simulations, we will briefly go through the fundamentals of the model we present. We leave the mathematical details of the implementation to the Methods section.

\subsection{Modeling context-dependent adaptation}
The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component occur in this order, and each component informs the ones that follow.

@@

We now briefly describe the three components of the model separately. For clarity, we start with the second component (i.e. motor adaptation), in order to introduce the terminology that we use throughout this work. For an overview, we show in \fref{fig:model} how these three components affect each other. The full mathematical description of the model can be found in the Methods.

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Schematic representation of the model. $\zeta$ represents the identity of the context as a categorical variable, $\vartheta$ are observations, $c_t$ are motor commands at trial $t$, $\phi$ are the parameters of the forward model and $s$ are the hidden states of the environment (e.g. the position of the hand). The arrows show flow of information, i.e. the dependencies in the update equations, as well as the flow of a single trial. Trial $t$ starts at the top when an observation $\vartheta_t$ is received, from which the current context is inferred ($q(\zeta)$), which informs motor adaptation ($q(\phi)$), which informs action selection ($p(c)$). The update equations can be found in the Methods section. After an action is selected, a new observation $\vartheta_{t+1}$ is received and the process is repeated. The posteriors over $\zeta$ and $\phi$ are used as priors at the next trial.}
\label{fig:model}
\end{figure}

\subsubsection{Motor adaptation}
It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. We used this principle and chose to make use of an exact Bayesian learner. To adapt to altered dynamics (e.g. a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model. These updates are done according to \eref{eqn:update-full}, using the difference between predictions made with the current parameters of the forward model, and the current observation. With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}, is calculated after each trial.

\subsubsection{Context inference}
To infer the adequate forward model to deploy, most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the forward models: the model that best predicts the current observation given the past is deemed to be the ``correct'' model, i.e. the model that best represents the current context. \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.

Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred as a probability distribution $q(\zeta)$ with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

\subsubsection{Action selection}
Action selection is affected by the current context via the context-dependent forward model that is used. In our model, action selection is made by building a distribution over available actions which is a weighted sum of the distributions given by the existing forward models, where the weights are obtained by the context inference component of the model.

In other words, a distribution $p(c_t)$ over all possible motor commands is created following \eref{eqn:dist-comm}, from which the next action (motor command $c_t$) can be sampled and carried out.

@@

In this section, we discuss savings and de-adaptation in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}. Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not, namely that savings are not immediate, but a relatively fast process that reflects context inference.

@@

We selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information available to participants. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, a colored light identified the current context, making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Savings and de-adaptation. Data from our simulations (left column) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (right column). Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown. In all panels, the blue line represents the adaptation displayed by the agent as a function of trial number. The black lines represent the optimal adaptation, i.e. the size of the true visuomotor rotation during the task. (A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O. Participants must adapt to the rotations during shooting movements, and the blue line can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation. When black and blue lines match, the participant perfectly counteracted the visuomotor rotation and reached the target. (B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced. O-A and A-O transitions (i.e. adaptation and de-adaptation) can be observed. Blue and black lines are as in A. Different blocks with the same adaptation are differentiated by the subscript (e.g. $A_1$ and $A_2$ have the same 20 degree visuomotor rotation). (C) Same experiment as (B) but with a 10 degree adaptation.} \label{fig:oh-2019}
\end{figure}

Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks. Switching between opposite adaptations was nevertheless immediate in simulations and experimental data. After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate. These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches. To expand on these results, we now turn to performance feedback and its effects on switching behavior.

\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model. Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$. From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation). In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up. As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C). This difference is explained by our model in terms of the size of the adaptation: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

@@

\subsubsection{Uncertainty over contexts affects action selection}
As with learning, our model selects actions (motor commands) based on context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eqn:dist-comm}).

@@

In our context-inference account, the asymmetry is caused by the baseline context, which have a non-zero probability $p(\zeta_O | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $\zeta_O$ becomes more likely (given that $\zeta_{3A}$ has been ruled out). Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}. Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data. However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.

\begin{figure}
\centering
\includegraphics{./figures/figure_3.png}
\caption{Motor error when switching back to a previously learned adaptation. Following the block structure from \cite{Davidson_Scaling_2004}, the first 40 trials of the switch from 3A (yellow) or -A (blue) to A is shown(A) Simulations from our model, with the task parameters from \cite{Davidson_Scaling_2004}, running eight simulations per group, each simulation representing one participant. On the y axis, the error is shown in arbitrary units, related to centimeters through a monotonically increasing transformation with the same origin. The shaded area is the SEM across simulated runs (participants). (B) Data adapted from figure 4 by \cite{Davidson_Scaling_2004}, showing the error in the same trials as (A).}
\label{fig:davidson-2004}
\end{figure}

@@

In \fref{fig:vaswani-2013}A, we show data simulated with our model, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}. The displayed adaptation is shown during the E trials for the three experimental groups in the experiment. It can be seen that group 1.3 (i.e. the participants who had learned in the -A/2 context in addition to A) more quickly recognized a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.

@@

In \fref{fig:vaswani-2013}C, the inference over context is shown for each group separately. Context inference works reliably until the error-clamp trials start, which do not correspond to any of the known contexts. This causes the agent to infer the combination of some of the known contexts that best fits the observations. Depending on the contexts previously learned by the agent: groups 1.1 and 1.4 display the same behavior, where the previous context (A and -A, respectively) slowly dwindles. These agents will slowly lower the force applied. In contrast, group 1.3 has learned the additional -A/2 context, which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero. Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4.

@@

\section{Discussion}
We presented a hierarchical model for motor adaptation based on Bayesian inference, which not only learns to adapt its motor outputs by observing errors, but also infers the context in which these errors were committed and updates the forward model associated with that context.

We generalized the inference over the best-fitting internal models to the concept of context inference, which includes all sensory information, in addition to feedback provided by the experimenters or the task. With this, we showed how this seemingly-simple generalization provides a parsimonious explanation for several experimental phenomena across many experiments in motor adaptation that, up to now, had assumed the existence of multiple mechanisms.

@@

Our choice of an exact Bayesian learner greatly simplified our simulations, but we expect the results we presented here to hold using alternative models for motor adaptation, such as Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}, as well as the recently-introduced COIN model \citep{Heald_Contextual_2021}. An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.

@@

\subsubsection{Reducing the effect of volatility on learning}
Experiments \citep{Herzfeld_memory_2014} have shown an effect of the volatility of the environment (i.e. unpredictable switching between contexts) on measured learning rate. Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment.

>>>>

\abstract{Abstract}
Humans have been shown to adapt their movements when sudden or gradual changes to the dynamics of the environment are introduced, a phenomenon called motor adaptation. If the change is reverted, the adaptation is also quickly reverted. Human are also able to adapt to multiple changes in dynamics presented separately, and to be able to switch between adapted movements on the fly. Such switching relies on contextual information which is often noisy or misleading, which affects the switch between adaptations. Recently, the COIN computational model for motor adaptation and context inference was introduced, which contains explicit components of context inference and Bayesian motor adaptation. This model was used to show the effects of context inference on learning rates across different experiments. We expanded on that work by using the COIN model to show that the effects of context inference on motor adaptation and control go further than previously shown. Here, we used the COIN model to simulate classical motor adaptation experiments from previous works and showed that context inference, and how it is affected by the presence and reliability of feedback, effect a host of behavioral phenomena that had so far required multiple hypothesized mechanisms, lacking a unified explanation. Concretely, we show that the reliability of direct contextual information, as well as noisy sensory feedback, typical of many experiments, effect measurable changes in switching-task behavior, as well as in action selection, that stem directly from imprecise context inference.

@@

Recently, a new computational model for context-dependant motor learning based on Bayesian inference was introduced \cite{Heald_Contextual_2021}, called COIN (for context inference). \cite{Heald_Contextual_2021} formalized context inference as a process that operates independently from motor learning, but is informed by it, establishing a loop whereby context inference also informs motor learning. With this model, \cite{Heald_Contextual_2021} showed that context inference causes the observed changes in the rate of motor learning un previous experiments \cite[e.g.][]{Herzfeld_Encoding_2018}.

In this work, we use the COIN model to show that the process of context inference underlies more behavioral phenomena than previously shown. we focused on the effects of uncertain contextual information on switching behavior, especially during so-called error-clamp trials, in which errors are forced to zero by experimenters. More specifically, we focused on the effects of perceptual noise, as well as feedback modalities, in context inference, which in turn affects behavior in ways that can be directly measured. We show that through context inference, switching behavior can display three main effects that have been previously attributed to hypothesized ad-hoc mechanisms: (1) The size of an adaptation dictates how quick and reliable switching between tasks is \cite{Oh_Minimizing_2019,Kim_Neural_2015}, which we explain in terms of the effects of perceptual noise on context inference. (2) Previously-learned adaptations can interfere with switching behavior \citep{Davidson_Scaling_2004}, which we explain in terms of uncertain context inference. (3) Training history (i.e. which adaptations have been learned and for how long) affects switching during error-clamp trials \citep{Vaswani_Decay_2013}, which we also attribute to uncertain context inference. To do this, we used the COIN model to simulate the experimental setups and the decision-making agents (i.e. participants) during those experiments.

@@

\section{Results}
Using the COIN model \citep{Heald_Contextual_2021}, we simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference. We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.

Before presenting these results, we make a brief description of the COIN model, leaving a more thorough explanation for the methods section. Additionally, we present simulations using the model that show the effects of contextual cues and perceptual noise on context inference, which pave the way for the simulated experiments that we show later on.

\subsection{Modeling context-dependent adaptation}
We focused on three main components of the COIN model: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component occur in this order, and each component informs the ones that follow.

@@

The COIN model contains, additionally to these three main components, components for learning new contexts (i.e. inferring the existence of a new context that had not been previously encountered by the agent), as well as the ability to infer subject-specific parameters such as a participant's assumed transition probabilities between contexts, which can differ from the real, hidden transition probabilities. In this work, we focus on switching behavior between known contexts, as well as in the perceptual aspect of context inference. Additionally, we do not fit the COIN model to participant data, instead focusing on showing the way the COIN model explains known experimental phenomena. For these reasons, we chose not to use these components of the COIN model, instead fixing both the transition probabilities between contexts and the number of contexts based on the real values during the experiments, which were made available to participants \todo{Check this}. Fixing these values, in turn, allowed us to simplify the generative model, internal to the agent, using conjugate priors, with which the agent can do exact Bayesian inference. The generative model we used here can be seen in \fref{fig:model}A, including the priors for both context inference and motor adaptation. For more details on these choices, see the Methods section.

\begin{figure}
\centering
\includegraphics[]{./figures/figure_1.png}
\caption{Schematic representation of the model and simulations. (A) Generative model. Clear circles represent the observables, i.e. motor commands $u_t$ and direct observations $y_t$ (e.g. cursor position). The real context $\zeta_t$ is not directly observable, but influences $y_t$. The dark rectangles represent the prior distributions for the inferred adaptation level $x_t$ (Normal-Gamma distribution) and the current context $\zeta_t$ (discrete distribution with known $\pi$; see Methods). At every trial, the context is inferred, then motor adaptation is carried out and finally a motor command is issued; the flow of this process is indicated by the gray arrow in the background, while black arrows show the direction of information flow. (B) Simulations obtained with the model in (A), using an experimental setup similar to that by \cite{Davidson_Scaling_2004}. The plots are organized as follows: each pair of plots represent context inference (bottom; p(ctx)) and response/adaptation (top). In the context inference plot, the y-axis represents the posterior probability of each context $p(\zeta_t = j$; the black line represents the baseline context (i.e. no adaptation), the green line represents the only adaptation to be learned during the simulated experiment. In the top plot (adaptation), the y-axis is the adaptation to be learned (e.g. the angle in a visuomotor rotation experiment), using the same color scheme (black for baseline). The purple line is the agent's response for the trial. The shaded areas represent the standard deviation around the mean, obtained across 50 simulated participants. Note, however, that the top plots are shown for a single run, for clarity (see main text). A total of 3x3 experiments were simulated, with low, medium and high levels of both cue uncertainty and observation noise. All plots share the same x-axis, which represents the trial number.}
\label{fig:model}
\end{figure}
\todo{Probably remove the top plots, as they are hard to explain and add very little to the paper}
While these priors are not mathematically equivalent to those used by \cite{Heald_Contextual_2021}, the prior samples they produce can be shown to be similar. For more details, see the supplementary materials. \todo{Do this}

\subsection{Contextual cues and feedback}
The behavioral phenomena which are the focus of this work can be explained as arising from the effects of contextual cues and sensory feedback provided to participants during the experiment. To show these effects, we first simulated a generic motor adaptation experiment similar to those performed by \cite{Davidson_Scaling_2004}, in which participants must adapt to a visuomotor rotation of the cursor on a screen in relation to the movement of a pointing device.

To show the effects that will be the base for the results presented below, we manipulated the presence and reliability of contextual cues, as well as the perceptual noise added to the cursor. In \fref{fig:model}B, a 3x3 grid of results can be seen, where each simulation in this grid is a combination of low, medium or high contextual cue reliability (where low reliability is equivalent to presenting no cues), and low, medium or high perceptual noise.

The first column of \fref{fig:model} shows that, in the presence of reliable contextual cues, context inference is correct, certain and fast to switch. However, as contextual cues become less reliable, switching between known contexts becomes slower, as seen in the posterior probabilities over contexts at and after trial 20 in the second and third columns. Furthermore, as perceptual noise increases, switching becomes not only slower, but also more uncertain, with individual agents incorrectly missing the switch entirely.

Additionally, the motor responses (shown in purple) can be seen to depend not only on the current estimate of the rotation angle (shown as the green line), but also the uncertainty over the current context. This is most evident in the high cue uncertainty and observation noise panel, in which the response varies substantially from the estimate of the angle (i.e. the purple line deviates from the green line), even when the real context has the highest estimated probability.

As we show below, these effects are at the heart of the behavioral phenomena observed in the experiments by \cite{Kim_Neural_2015}, \cite{Oh_Minimizing_2019}, \cite{Davidson_Scaling_2004}, and \cite{Vaswani_Decay_2013}, which we directly simulate in this work, alongside others that we discuss in the Discussion section.

@@

In this section, we discuss savings and de-adaptation in terms of switching between contexts. We show that through context inference and how its affected by contextual cues and obsevation noise, savings are not immediate, but a relatively fast process that reflects context inference.

@@

We selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information available to participants. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, a colored light identified the current context, making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate.

In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts in the center panel, i.e. when the solid gray line (representing the true context) switches between 40, 0 and -40; in both panels (second and third), the thick black line, representing the agent's adaptation, quickly follows these switches.

\begin{figure}
\centering
\includegraphics{./figures/figure_2.png}
\caption{Savings and de-adaptation. Data from our simulations (first and second columns) compared to data adapted from figure 2A by \cite{Kim_Neural_2015} and figure 4A from \cite{Oh_Minimizing_2019} (last column). Experimental data was, in all three experiments, averaged across all participants; in our simulations, a single simulation is shown. In the first column, the simulated context over inference is represented by the posterior probabilities over all available contexts. Each color (black, blue, green) represents a different context, with black always representing the baseline (i.e. no adaptation). Vertical, dashed lines represent switches in the real context. In the second and third columns, the black line represents the adaptation (i.e. response) displayed by the agent as a function of trial number. The thin gray lines represent the optimal adaptation, i.e. the size of the true visuomotor rotation during the task. In all panels, the x-axis represents trial number, from 0 to 300. (A) Experiment by \cite{Kim_Neural_2015} with two visuomotor rotations (-40 and 40 degrees), in addition to baseline, with all possible transitions between A, (-A) and O. Participants must adapt to the rotations during shooting movements, and the participant's response can be understood as the rotation of the participants' shooting movements to compensate for the task's rotation. For clarity, both experimental and simulated results are shown only up to trial 300, of the original 600. (B) An experiment by \cite{Oh_Minimizing_2019} similar to A, but the contexts were not cued and only one context, with an adaptation of 20 degrees, is introduced. O-A and A-O transitions (i.e. adaptation and de-adaptation) can be observed. Blue and black lines are as in A. (C) Same experiment as (B) but with a 10 degree adaptation.}
\label{fig:oh-2019}
\end{figure}

Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by their responses not being on par with the true rotation angle. This undershooting of responses (i.e. not doing the 40 degree rotation) happens despite participants being able to immediately and with high certainty identify the real context, as shown in the first-column panel. These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches. To expand on these results, we now turn to feedback and its effects on switching behavior.

\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model. Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A$. In subsequent $A - O$ transitions, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in their responses (black line in the last two columns of \fref{fig:oh-2019}B) closely following the switches in the black line (true rotation). In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up. As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C). This difference is explained in our simulations in terms of the size of the adaptation in relation to observation noise: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

@@

\subsubsection{Uncertainty over contexts affects action selection}
As with learning, action selection is affected by context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eqn:dist-comm}).

@@

In the context-inference account, the asymmetry is caused by the existance of the baseline context, which has a non-zero probability $p(\zeta_O | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $\zeta_O$ becomes more likely (given that $\zeta_{3A}$ has been ruled out). Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}A, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}.

\begin{figure}
\centering
\includegraphics{./figures/figure_3.png}
\caption{Motor error when switching back to a previously-learned adaptation. (A) Experimental results from \cite{Davidson_Scaling_2004} and simulations with the COIN model are shown. In the first column, each panel represents context inference for one group of participants (top: group A; bottom: group 3A), with each line representing the posterior probability of a context (black for the baseline O). The second column represents the error made by our simulated agent after returning to the previously-learned context, with blue and yellow representing groups -A and 3A, respectively. The last panel represents the same data, from the \cite{Davidson_Scaling_2004} experiment. All panels share the x-axis, representing the number of trials elapsed since the switch to the new context. All simulations were executed eight times per group, following the number of participants in the experiments; shaded areas represent the standard deviation across these eight simulations. (B) Simulated results for an experiment similar to \cite{Davidson_Scaling_2004}, but changing the contexts seen by the two groups from -A to -2A for the first group, and from 3A to 4A for the second group. The panels follow the same structure as (A), without the last panel for experimental results. (C) Simulations for an experiment in which the baseline context has been removed altogether.}
\label{fig:davidson-2004}
\end{figure}

To confirm our hypothesis, we simulated variants of the experiments in which the context-inference account predicts that the difference between groups disappears. First, in \fref{fig:davidson-2004}B, we simulated an experiment in which the contexts have more extreme adaptations, making them more different from baseline than in the \cite{Davidson_Scaling_2004} experiments. To do this, one group adapts in a $O-A-(-2A)$ paradigm, while the other group adapts in a $0-A-4A$ paradigm. As in the original experiment, the second contexts ($-2A$ for one group, $4A$ for the other) are equally spaced from the first context. However, given the larger distance from baseline, the baseline context has a much lower attribution when the switch occurs compared to the original experiment. This change makes both simulated groups infer the correct context in much more similar times, making the difference between their errors be much smaller compared to the original experiment. Furthermore, we simulated an experiment with an identical structure to that of \cite{Davidson_Scaling_2004}, but eliminated the baseline context from the agent. The results can be seen in \fref{fig:davidson-2004}C, where the switches between contexts are made identically by the two groups. We discuss possible implementations of such an experiment in the Discussion section, alongside with other predictions made by the model. \todo{Promises, promises}

@@

In \fref{fig:vaswani-2013}A, we show data simulated with the COIN, following the parameters of the experiment by \cite{Vaswani_Decay_2013}, and in \fref{fig:vaswani-2013}B we show the experimental plots adapted from \cite{Vaswani_Decay_2013}. The displayed adaptation is shown during the E trials for the three experimental groups in the experiment. It can be seen that group 1.3 (i.e. the participants who had learned in the -A/2 context in addition to A) more quickly recognized a change in context and lowers the force applied on the mechanical handle, as can be seen in the experimental data.

@@

In \fref{fig:vaswani-2013}C, the inference over context is shown for each group separately. Context inference works reliably until the error-clamp trials start, which do not correspond to any of the known contexts. This causes the agent to infer the combination of some of the known contexts that best fits the observations. Depending on the contexts previously learned by the agent: groups 1.1 and 1.4 display the same behavior, where the previous context (A and -A, respectively) slowly dwindles. These agents will slowly lower the force applied. In contrast, group 1.3 has learned the additional -A/2 context, which has a non-zero posterior probability during E trials, pushing the agent's adaptation force more quickly towards zero. Group 1.2 behaves similarly to 1.1, with the exception that the baseline context, which was recently seen, plays a bigger role during E trials, making the agent reduce its force during E trials slightly more quickly than groups 1.1 and 1.4. \todo{Danger! Danger! A keen observer might point out that the model might want to learn a new context, as COIN does. What to do? What to do?}

@@

\section{Discussion}
We expand on previous works \cite{Heald_Contextual_2021,Oh_Minimizing_2019} that introduced the idea that context inference is a process that informs and is informed by motor adaptation by showing that it explains behavioral phenomena that had previously required different specific, ad-hoc mechanisms outside of contextual motor adaptation.

@@

We used the recently-introduced COIN model \cite{Heald_Contextual_2021}, with a different choice of prior distributions that greatly simplified our simulations without sacrificing relevant parts of the original specification. While we focused on the Bayesian motor adaptation of the COIN model, we expect the results we presented here to hold using alternative models for motor adaptation, such as other Kalman filter-based learners \cite[e.g.][]{Oh_Minimizing_2019,Baddeley_System_2003} and other Bayesian implementations \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004}. An exception is linear learners \cite[e.g.][]{Smith_Interacting_2006,Forano_Timescales_2020,Lee_Dual_2009}, as these models do not incorporate uncertainty in the estimate of the parameters of the forward models.

@@

Note that the original COIN specification includes a component to learn new contexts. However, this component works exclusively by creating new contexts in which the forward models take the same form but have different parameter values. It is unclear whether the model can be adapted to create contexts in an online fashion that operate in an essentially different manner, as is the case of error-clamp trials, in which participants' responses do not affect the outcome and motor commands are issued based on criteria not directly related to the goal of the task (e.g. energy minimization or comfort maximization).

\subsubsection{Reducing the effect of volatility on learning}
Experiments \citep{Herzfeld_memory_2014} have shown an effect of the volatility of the environment (i.e. unpredictable switching between contexts) on measured learning rate. Our model predicts that this effect would be greatly reduced if reliable contextual cues were introduced: if a participant can infer the context of the current trial before any decision or observation has occurred, the learning rate would not be affected by the volatility of the environment. \cite{Heald_Contextual_2021} showed that the COIN model explains this change in terms of the Kalman gain and the posterior probability over contexts (see their equation 14) \todo{Check this in more detail}.
